---
title: "Machine Learning: personal activity"
author: "Xiaojing HU"
date: "Tuesday, June 16, 2015"
output: html_document
---

# Load and prepare the data
```{r load data, results=FALSE, cache=TRUE}
data <- read.csv("pml-training.csv")
dim(data)
table(data$classe) #manner to be predicted
```

Using summary(data), learning about variables.
Some variables with too many NAs or Blank, and first 7 variables about the number(X), user, time and windows, are also irrelevant.
Removing these variables from predictors.

```{r remove variables, cache=TRUE}
v <- apply(is.na(data)|data=="", 2, sum)
sum(v==0); sum(v>19000)
data <- subset(data[,v==0]); data <- data[-c(1:7)]
dim(data)
```

# Explore the data
Now we have 53 variables, combination of:
- 6 different ways: roll, pitch, yaw, gyros, accel and magnet
- 4 accelerometer locations: belt, arm, dumbbell, forearm

Now exploring variables with some plots.

```{r plot, echo=FALSE, warning=FALSE}
library(ggplot2)
qplot(gyros_belt_x, gyros_belt_y, data=data, color=classe)
qplot(gyros_dumbbell_y, data=data, fill=classe, geom="density", xlim=c(0:1))
```

# Machine learning
## Split data to training & testing sets for cross validation
```{r prepare data sets, cache=TRUE}
set.seed(23123)
library(caret)
inTrain <- createDataPartition(y=data$classe, p=0.6, list=FALSE)
train <- data[inTrain,]; test <- data[-inTrain,]
```

## Classification Model
It's a question of classification.

Trying to use classification tree, random forest, model based prediction nad moosting. With default setting, method="lda" is best for its accuracy and speed.

The Accuracy is about 0.70, so the out of sample error is expected 0.30.

```{r model, cache=TRUE}
modFit <- train(classe~., method="lda", data=train)
modFit$result
pred <- predict(modFit, test)
table(pred, test$classe)
sum(pred==test$classe)/dim(test)[1] # Testing Set's Accuracy
```
The accuracy of testing set is a little better than that of training set.

It's not so good. But is as expected.

# Machine learning algorithm apply to Test data
For submission of predictions, use the model to original Test data.
```{r answer, cache=TRUE}
testing <- read.csv("pml-testing.csv")
answer <- predict(modFit, testing); answer
```


